|

### 2022-08-22 appendix/appendix.Rmd
update short article and appendix

In our first experiment, we analyze how respondents justify their preferences
towards extralegal violence. We assess the impact of three factors that have
been cited as major drivers of lynchings: 1) police ineffectiveness; 2) slow
criminal justice; 3) demand for harsher punishment for criminals. Below, we
discuss them in further detail.

Research shows that police ineffectiveness frequently appears as a strong
predictor of vigilantism [@cruz2019determinants; @garcia2019anger]. The direct
result of the weakness of police institutions is that citizens decide to take
criminal matters "into their own hands", thus persecuting and punishing the
criminals by themselves. A recent statistic indicates that the police solves
only 10% of the homicides in Brazil, which lends support to the link between
weak law enforcement and lynchings [@pearson2018latam].

Another possible determinant of lynching support is lack of trust in the
justice system [@godoy2004justice; @smith2019contradictions]. This is often due
to long criminal proceedings, which cause significant anxiety for the victims.
In Brazil, the penal code allows the accused to appeal each decision several
times, so it can take decades before a criminal case is closed
[@sousa2005utilizaccao]. In this respect, citizens do not believe that, even if
the criminal is put to trial, he/she will be punished in a timely matter. Note
that although the police is technically part of the criminal justice system, we
analyze the two institutions separately in our experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their crimes. In
particular, we intend to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
addresses not the efficiency of the institutions, but their legitimacy
[@nivette2016institutional]. In fact, Brazilians are often vocal about their
preference for repressive legal punishment. In a recent article in _The Wall
Street Journal_, a bar owner justified the lynching of the local thug who
killed his son by saying that "_even if he had been put behind bars for 100
years it wouldn't have been enough to pay for all his crimes_"
[@pearson2018latam]. We hypothesize that many Brazilians also share this view.

The experiment consists of three treatment conditions and one control group.
Respondents read an excerpt of a news article describing a real lynching case.
We have slightly edited the original text so that respondents have no prior
knowledge of the crime.[^jornal] The vignette for the control group includes no
information about the reasons behind the lynching. We ask respondents to show
their level of lynching support using a 0-100 slider, where 0 means no support
and 100 means full support. Respondents in each of the three treatment arms
read the same piece, but with one additional sentence explaining the
motivations behind the lynching. The vignettes are as follows:

[^jornal]: The original article is available at the following address:
<https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso>.
Access: August 2020.

* *Control group*: A man was lynched last Friday in Jundiaí, São Paulo.
  According to the neighbours, he tried to break into a house but was
  immobilised and beaten by members of the community.[^control-exp01]

[^control-exp01]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade.

* *Treatment 01 - Police ineffectiveness*: A man was lynched last Friday in
  Jundiaí, São Paulo. According to the neighbours, he tried to break into a
  house but was immobilised and beaten by members of the community. **One of
  the residents who took part in the lynching said they had beaten the suspect
  because "the police never patrols the area"**.[^treat01-exp01]

[^treat01-exp01]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a polícia
nunca patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow and the perpetrator is on
  the street until the case is heard"**.[^treat02-exp01]

[^treat02-exp01]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta e os criminosos ficam soltos até o julgamento"**.

* *Treatment 03 - Demand for harsher legal punishment*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial punishment is not harsh
  enough"**.[^treat03-exp01]

[^treat03-exp01]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a punição
da justiça não é dura o suficiente"**.

Before each vignette, respondents read the following text:

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask you one question about it.[^vig01-exp01]

[^vig01-exp01]: In Portuguese: Uma notícia será apresentada para você. Por
favor, leia a notícia com atenção. Após você ler o artigo, faremos uma pergunta
sobre ele.

After the vignette, respondents were presented with this question:

* Do you think that the lynching was justified? Please use the slider below
  to indicate your opinion. For disagreement, use 0 to 49; for agreement,
  use 51 to 100. Please use 50 if you neither agree nor
  disagree.[^vig02-exp01]

[^vig02-exp01]: In Portuguese: Você acha que o linchamento foi correto? Por
favor, use a barra abaixo para indicar sua opinião. Para discordar, use de 0 a
49; para concordar, use de 51 a 100. Por favor, use 50 para não concordar nem
discordar.

Table \ref{tab:exp01main} summarizes our results. As we see, none of the
treatment effects are statistically significant at conventional levels. This
suggests that institutional factors do not affect lynching support in our
sample.

\footnotesize
```{r exp01-main, warning=F, message=F, error=F, cache=T, fig.pos="H", results='asis'}
df_exp01 <- df1 %>%
  mutate(exp01_outcomes = coalesce(exp01_control, exp01_police,
                                   exp01_slow_justice, exp01_small_punishment),
         exp01_any_treat = case_when(!is.na(exp01_control) ~ "0",
                                     !is.na(exp01_police) ~ "1",
                                     !is.na(exp01_slow_justice) ~ "1",
                                     !is.na(exp01_small_punishment) ~ "1",
                                     TRUE ~ NA_character_),
         exp01_police_treat = case_when(!is.na(exp01_control) ~ "0",
                                        !is.na(exp01_police) ~ "1"),
         exp01_slow_justice_treat = case_when(!is.na(exp01_control) ~ "0",
                                              !is.na(exp01_slow_justice) ~ "1"),
         exp01_small_punishment_treat = case_when(!is.na(exp01_control) ~ "0",
                                                  !is.na(exp01_small_punishment) ~ "1"))

m1 <- lm(exp01_outcomes ~ exp01_police_treat, data = df_exp01)
m2 <- lm(exp01_outcomes ~ exp01_slow_justice_treat, data = df_exp01)
m3 <- lm(exp01_outcomes ~ exp01_small_punishment_treat, data = df_exp01)
m4 <- lm(exp01_outcomes ~ exp01_any_treat, data = df_exp01)

stargazer(m1, m2, m3, m4, se = starprep(m1, m2, m3, m4), p = starprep(m1, m2, m3, m4, stat = "p.value"),
          header = FALSE, align = TRUE, label = "tab:exp01main", title = "Experiment 01 -- Main Results",
          style = "apsr", dep.var.labels = "\\textbf{Lynching Support}\\vspace{.5cm}",
          covariate.labels = c("Police does not patrol area", "Justice too slow",
                               "Punishment not harsh enough", "Combined treatments"),
          column.sep.width = "3pt", notes = "Robust standard errors in parentheses.",
          keep.stat = "n", no.space = TRUE)
```
\normalsize

Here we estimate the effect of gender, race, and political ideology on lynching
support. We find that males are more likely to tolerate lynchings, whereas left
and center-left voters oppose lynchings more strongly than centrists. In our
second model, we find no effect for race, and in the last column we see that
white respondents are less supportive of lynchings.

\footnotesize
```{r exp01-baseline, warning=F, message=F, error=F, cache=T, fig.pos="H", results='asis'}
df_exp01_group <- df_exp01 %>%
  filter(gender == c("Female", "Male"),
         race %in% c("Asian", "Black", "Mixed Race", "White"),
         ideology %in% c("Center", "Center-Left", "Center-Right", "Left", "Right")) %>%
  mutate(race = fct_relevel(race, "Black"), ideology = fct_relevel(ideology, "Center"))

df_exp01_gender <- df_exp01 %>% filter(gender == c("Female", "Male"))
df_exp01_race <- df_exp01 %>% filter(race %in% c("Asian", "Black", "Mixed Race", "White")) %>%
  mutate(race = fct_relevel(race, "Black"))
df_exp01_ideology <- df_exp01 %>%
         filter(ideology %in% c("Center", "Center-Left", "Center-Right", "Left", "Right")) %>%
         mutate(ideology = fct_relevel(ideology, "Center"))

m1 <- lm(exp01_outcomes ~ gender, data = df_exp01_gender)
m2 <- lm(exp01_outcomes ~ race, data = df_exp01_race)
m3 <- lm(exp01_outcomes ~ ideology, data = df_exp01_ideology)
m4 <- lm(exp01_outcomes ~ gender + race + ideology, data = df_exp01_group)

stargazer(m1, m2, m3, m4, se = starprep(m1, m2, m3, m4), p = starprep(m1, m2, m3, m4, stat = "p.value"),
          header = FALSE, align = TRUE, label = "tab:exp01baseline",
          title = "Experiment 01 -- Determinants of Baseline Levels of
          Lynching Support", style = "apsr", dep.var.labels = "\\textbf{Lynching Support}\\vspace{.5cm}",
          covariate.labels = c("Male", "Asian", "Mixed Race", "White",
                               "Left", "Center-Left", "Center-Right", "Right"),
          column.sep.width = "3pt", notes = "Robust standard errors in parentheses.",
          keep.stat = "n", no.space = TRUE)
```
\normalsize

We estimate heterogeneous effects for our experiment using Bayesian Additive
Regression Trees (BART) [@chipman2010bart; @hill2011bayesian]. BART methods
allow users to detect non-linear interactions and are insensitive to the choice
of tuning parameters, so they are well-suited to analyze observational and
experimental data [@green2012modeling]. More specifically, we employ the
`bartCause` package, which is designed to estimate causal inference models.

### Treatment 01: Police Ineffectiveness

Our results indicate that the effects for this treatment arm is null in every
model specification. We find no evidence of heterogeneous effects.

\footnotesize
```{r exp01-bart01, warning=F, error=F, message=F, cache=T}
df_exp01_het <- df_exp01 %>%
  filter(gender %in% c("Female", "Male")) %>%
  mutate(race = fct_relevel(race, "White", "Black", "Mixed Race", "Asian",
                            "Indigenous"),
         education = fct_relevel(education, "Primary School", "Secondary School",
                                  "High School", "College", "Graduate School"),
         views_police = fct_relevel(views_police, "Regular", "Very Good", "Good",
                                    "Bad", "Very Bad"),
         views_justice = fct_relevel(views_justice, "Regular", "Very Good", "Good",
                                     "Bad", "Very Bad"),
         ideology = fct_relevel(ideology, "Center", "Left", "Center-Left",
                                "Center-Right", "Right", "Don't Know", "Rather Not Say"),
         household_income = fct_relevel(household_income, "Up to R$1,000", "From R$1,001 to R$2,000",
                                         "From R$2,001 to R$3,000", "From R$3,001 to R$5,000",
                                         "From R$5,001 to R$10,000", "From R$10,001 to R$20,000",
                                         "Above R$20,000"),
         previous_victim_dummy = fct_relevel(previous_victim_dummy, "Yes", "No"),
         death_penalty = fct_relevel(death_penalty, "Yes", "No"),
         age2 = case_when(age >= 18 & age <= 34 ~ "18-34", age >= 35 & age <= 54 ~ "35-54",
                          age >= 55 ~ "55 plus", TRUE ~ as.character(age)))

df_exp01_police <- df_exp01_het %>%
  mutate(exp01_police_treat = as.numeric(exp01_police_treat)) %>%
  drop_na(exp01_police_treat)

# Gender
summary(bartc(exp01_outcomes, exp01_police_treat, gender,
              group.by = gender, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Age
summary(bartc(exp01_outcomes, exp01_police_treat, age2,
              group.by = age2, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Race
summary(bartc(exp01_outcomes, exp01_police_treat, race,
              group.by = race, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Education
summary(bartc(exp01_outcomes, exp01_police_treat, education,
              group.by = education, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Household Income
summary(bartc(exp01_outcomes, exp01_police_treat, household_income,
              group.by = household_income, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Political Ideology
summary(bartc(exp01_outcomes, exp01_police_treat, ideology,
              group.by = ideology, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Support for Death Penalty
summary(bartc(exp01_outcomes, exp01_police_treat, death_penalty,
              group.by = death_penalty, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Previous Victimization
df_exp01_police2 <- df_exp01_police %>% drop_na(previous_victim_dummy)
summary(bartc(exp01_outcomes, exp01_police_treat, previous_victim_dummy,
              group.by = previous_victim_dummy, group.effects = TRUE, data = df_exp01_police2,
              n.chains = 5L, seed = 144))

# Opinion on the Police
summary(bartc(exp01_outcomes, exp01_police_treat, views_police,
              group.by = views_police, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))

# Opinion on the Judicial System
summary(bartc(exp01_outcomes, exp01_police_treat, views_justice,
              group.by = views_police, group.effects = TRUE, data = df_exp01_police,
              n.chains = 5L, seed = 144))
```
\normalsize

### Treatment 02: Criminal Justice Ineffectiveness

We do not find any evidence of heterogeneous treatment effects in this
condition either.

\footnotesize
```{r exp01-bart02, warning=F, error=F, message=F, cache=T}
# Remove "Indigenous" and "Other" from `race` as they only
# have 1 observation each. The model cannot be estimated otherwise.
df_exp01_slow_justice <- df_exp01_het %>%
  filter(race %in% c("White", "Asian", "Black", "Mixed Race", "Rather Not Say")) %>%
  mutate(exp01_slow_justice_treat = as.numeric(exp01_slow_justice_treat)) %>%
  drop_na(exp01_slow_justice_treat)

# Gender
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, gender,
              group.by = gender, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Age
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, age2,
              group.by = age2, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Race
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, race,
              group.by = race, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Education
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, education,
              group.by = education, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Household Income
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, household_income,
              group.by = household_income, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Political Ideology
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, ideology,
              group.by = ideology, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Support for Death Penalty
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, death_penalty,
              group.by = death_penalty, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Previous Victimization
df_exp01_slow_justice2 <- df_exp01_slow_justice %>% drop_na(previous_victim_dummy)
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, previous_victim_dummy,
              group.by = previous_victim_dummy, group.effects = TRUE, data = df_exp01_slow_justice2,
              n.chains = 5L, seed = 144))

# Opinion on the Police
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, views_police,
              group.by = views_police, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))

# Opinion on the Judicial System
summary(bartc(exp01_outcomes, exp01_slow_justice_treat, views_justice,
              group.by = views_police, group.effects = TRUE, data = df_exp01_slow_justice,
              n.chains = 5L, seed = 144))
```
\normalsize

### Treatment 03: Demand for Harsher Legal Punishment

As in the last two series of models, our results show no evidence of
heterogeneous treatment effects.

\footnotesize
```{r exp01-bart03, warning=F, error=F, message=F, cache=T}
df_exp01_small_punishment <- df_exp01_het %>%
  mutate(exp01_small_punishment_treat = as.numeric(exp01_small_punishment_treat)) %>%
  drop_na(exp01_small_punishment_treat)

# Gender
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, gender,
              group.by = gender, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Age
df_exp01_small_punishment2 <- df_exp01_small_punishment %>% drop_na(age2)
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, age2,
              group.by = age2, group.effects = TRUE, data = df_exp01_small_punishment2,
              n.chains = 5L, seed = 144))

# Race
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, race,
              group.by = race, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Education
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, education,
              group.by = education, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Household Income
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, household_income,
              group.by = household_income, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Political Ideology
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, ideology,
              group.by = ideology, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Support for Death Penalty
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, death_penalty,
              group.by = death_penalty, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Previous Victimization
df_exp01_small_punishment2 <- df_exp01_small_punishment %>% drop_na(previous_victim_dummy)
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, previous_victim_dummy,
              group.by = previous_victim_dummy, group.effects = TRUE, data = df_exp01_small_punishment2,
              n.chains = 5L, seed = 144))

# Opinion on the Police
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, views_police,
              group.by = views_police, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))

# Opinion on the Judicial System
summary(bartc(exp01_outcomes, exp01_small_punishment_treat, views_justice,
              group.by = views_police, group.effects = TRUE, data = df_exp01_small_punishment,
              n.chains = 5L, seed = 144))
```
\normalsize

In our second experiment, we present five pairs of criminal profiles to
respondents. Each profile consists of eight attributes: 1) gender of the crime
perpetrator; 2) age of the crime perpetrator; 3) race of the crime perpetrator;
4) residency of crime perpetrator; 5) offense; 6) gender of the victim of the
motivating crime; 7) age of the victim of the motivating crime; 8) lynching
perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

@@

# Experiment 03

@@

Our last experiment measures the effect of information provision on attitudes
about lynching. In particular, we test whether reminding respondents
about the legal and social consequences of vigilante justice reduces the
subjects' level of support for such practice.

@@

As we did in the first experiment, we evaluate how individual characteristics
impact lynching support. The results are similar to those presented in table
\ref{tab:exp01baseline}, with the exception that the coefficient for white
respondents is negative in two estimations, and the coefficient for male does
not reach statistical significance in the last model ($p$-value = 0.11).

>>>>

In our first experiment, we present five pairs of criminal profiles to
respondents. Each profile consists of eight attributes: 1) gender of the crime
perpetrator; 2) age of the crime perpetrator; 3) race of the crime perpetrator;
4) residency of crime perpetrator; 5) offense; 6) gender of the victim of the
motivating crime; 7) age of the victim of the motivating crime; 8) lynching
perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

@@

@@

Our second experiment measures the effect of information provision on attitudes
about lynching. In particular, we test whether reminding respondents
about the legal and social consequences of vigilante justice reduces the
subjects' level of support for such practice.

@@

We also evaluate how individual characteristics
impact lynching support. We find that the coefficient for white
respondents is negative in two estimations, and the coefficient for male does
not reach statistical significance in the last model ($p$-value = 0.11).
Political ideology is strongly correlated with support for lynchings.
